Response to reviewers

### 2025-01-21 Manuscript/ecography_reponse_letter_2.tex
added raw values
Update ecography_reponse_letter_2.tex
up response

% TODO: double check the definition and make sure it aligns with the correct one. Say we fucked up explanation because of the negative effect (i.e. removing X is the same as keeping not X).

@@

% TODO: re double checking how the scaling works and what it does to the data!

@@

% TODO: redo the supplementary plots but with different scale bars for every plot (and don't do the scaling)

>>>>

We believe the confusion stems from what probability we are talking here.
In a way here we approximate both competitive exclusion (i.e. increased ``extinction'' in densely populated regions) the same way as increased ``survival'' in sparsely populated regions.
The reviewer is true that these two mechanisms are fundamentally different in ecology but we believe that for our approximations, they result in the same patterns: extinction in dense region (due to high competition) and survival in sparse regions (due to low competition).

The formula illustrates how we measure the probability of being sampled which is decreasing in densely populated regions.
This probability of being sampled reflects the probability of ``surviving'', i.e. of not being removed from the simulation.
We've now added this precision throughout the explanation:

\textit{(or, inversely, observations are more likely to be ``preserved''/``surviving'' in less dense regions)}. Lines @@@

@@

% TODO: re double checking how the scaling works and what it does to the data!
% USE SES instead?

@@

We have now updated all these supplementary figures to show the raw results with the scale bars indicating the range of values for each metric.

### 2025-01-21 Manuscript/main_ecography_R2.tex
up response

\item Increasing evenness: by resampling the proportion of data (i.e. 20\%, 40\%, 60\% or 80\%) but with skewed resampling probabilities.
In brief, this algorithm reduces the probability of resampling observations in regions of the trait space that have many observations and increases it in regions that have few observations.
This is done by selecting $B$ discrete categories to summarise the distribution - i.e. bandwidths - using Silverman's ``rule of thumb'' \citep[\texttt{bw.nrd0} function in \texttt{R};][]{silverman1986density}.
Each discrete category $B$ has an observed probability of sampling of $b$ (the proportion of observations in the category) and each observation then gets a probability of resampling scaling with the intensity of the stressor (the proportion of what to remove $i$) of $i \times (1-b)^{p}$. 
Where $p$ is a factor increasing the scaling power of the algorithm (here we arbitrarily used $p=3$ which was a relatively low scaling value that still resulted in visible changes observable in 2D).
In other words, observations in categories with a low density ($b<0.5$) were more likely to be resampled and the ones in categories with high density ($b>0.5$) were less likely to be resampled.
This approximates our \textbf{competition mechanism}, where observations in dense regions of the trait space are more likely to be removed than in sparse regions of the trait space.
We chose this algorithm to approximate a stressor that could increase the probability of extinction for observations that share traits combination.
This could approximate the fact that observations that have similar trait combinations are more likely to go extinct due to competition (e.g. using the same resources). 
\end{itemize}

>>>>

\item Increasing evenness: by resampling the proportion of data (i.e. 20\%, 40\%, 60\% or 80\%) but with skewed resampling probabilities.
In brief, this algorithm reduces the probability of resampling observations in regions of the trait space that have many observations and increases it in regions that have few observations.
This is done by selecting $B$ discrete categories to summarise the distribution - i.e. bandwidths - using Silverman's ``rule of thumb'' \citep[\texttt{bw.nrd0} function in \texttt{R};][]{silverman1986density}.
Each discrete category $B$ has an observed probability of sampling of $b$ (the proportion of observations in the category) and each observation then gets a probability of resampling scaling with the intensity of the stressor (the proportion of what to remove $i$) of $i \times (1-b)^{p}$. 
Where $p$ is a factor increasing the scaling power of the algorithm (here we arbitrarily used $p=3$ which was a relatively low scaling value that still resulted in visible changes observable in 2D).
In other words, observations in categories with a low density ($b<0.5$) were more likely to be resampled and the ones in categories with high density ($b>0.5$) were less likely to be resampled.
This approximates our \textbf{competition mechanism}, where observations in dense regions of the trait space are more likely to be removed than in sparse regions of the trait space (or, inversely, observations are more likely to be ``preserved''/``surviving'' in less dense regions).
We chose this algorithm to approximate a stressor that could increase the probability of extinction for observations that share traits combination.
This could approximate the fact that observations that have similar trait combinations are more likely to go extinct due to competition (e.g. using the same resources). 
\end{itemize}
